Security vs. Maintainability: \\ Fixing Vulnerabilities Obfuscates your Code

### 2019-01-29 paper/ICPC19/icpc19.tex
abstract fixed, resuls finished, discussion introduced + rq brieflyanswered + threats to validity finished

\begin{abstract}
  Security is a crucial non-functionality requirement for software applications.
  However, building secure software is far from trivial as developers lack both
  the knowledge and tools to effectively address this concern. In this paper, we
  study the impact of changes to improve security on the maintainability of several
  open source applications. Using a dataset containing 607 security-oriented
  commits, we measure maintainability --- as computed by the Software Improvement
  Group's web-based source code analysis service \emph{Better Code Hub} (BCH) ---
  before and after the security refactoring. Results show that making software
  more secure comes at a cost on maintainability. This is particularly evident
  in refactorings to deal with \textit{Broken Authentication} and \textit{Cross-Site Request Forgery} attacks.
\end{abstract}

@@

The impact of security refactorings on the maintainability of open-source software per group of security flaws is presented in Figure \ref{fig:pat}. The results of the Wilcoxon signed-rank test by pattern returned statistical evidence which supports the maintainability of open-source software decreases for 4 different patterns: Broken Autentication (p = 0.018), Denial-of-Service (p = 0.029), Memory Leak (p = 0.009) and Miscellaneous (p = $3.07$x$10^{-4}$). The results also returned statistical significance that evidence that maintainability is more likely to not suffer differences for open-source software after removing Components with Known Vulnerabilities (p = $7.96$x$10^{-4}$) and fixing Cross-Site Scripting Vulnerabilities (p = 0.001). Injection and Cross-Site Request Forgery produced more cases where the refactorings had a negative impact in the maintainability. However, the results obtained for these patterns are not statistically significant.

For the Broken Authentication pattern, 55.3\% of the refactorings (21) had a negative impact in the maintainability, 28.95\% (11) improved maintainability and 15.8\% (6) of the changes did not add any impact in the maintainability of open-source software.

%    neg  pos  nul
% ml   41   28   11
% injec   32   31   26
% xss   27   24   63
% misc   84   62   49
% dos   19   12    9
% csrf   17    9    4
% auth   21   11    6
% ucwkv    3    1   17

%        neg       pos       nul
% ml  0.512500  0.350000  0.137500
% injec  0.359551  0.348315  0.292135
% xss  0.236842  0.210526  0.552632
% misc  0.430769  0.317949  0.251282
% dos  0.475000  0.300000  0.225000
% csrf  0.566667  0.300000  0.133333
% auth  0.552632  0.289474  0.157895
% ucwkv  0.142857  0.047619  0.809524

@@

@@

@@

\section{Threats to Validity}\label{sec:threats}
%
In this section, we discuss potential threats to the validity of our work.
%
\subsection{Construct}
%
We use a metric derived from the reports provided by BCH as a construct of
software maintainability. Different projects and contexts may require different
maintainability standards. However, BCH uses a representative benchmark of
closed and open source software systems to compute the thresholds used in each
maintainability guideline\cite{Visser:2016:OREILLY,baggen2012standardized}.
In addition, the benchmark is updated every year~\cite{Visser:2016:OREILLY}.

The lifespan of a given energy commit is an approximation based on a single
change in one of the lines of code submitted by that commit. Although changes in
whitespace are not considered, we did not perform any evaluation on the type of
change that is performed. Some changes may be trivial while others may have
required more efforts from developers. Related work has proposed approaches to
automatically classify changes (e.g., bug fixes\cite{da2017framework}, and
refactorings~\cite{pascarella2018self}), but we have not explored them.
Nevertheless, we understand that the differences in the lifespan between our
dataset of energy changes and the baseline provide interesting leads for
discussion and new research directions.

@@

We resort to a set of security changes that were collected from previous
work~\cite{Reis:2017:IJSSE}. These dataset was collected using the commit
message provided by developers to classify a given commit as a security change.
We understand that this approach discards security changes that did not have an
explicit commit message describing them as such.

In the case of baseline commits, they were collected by randomly select a commit
in the same project that a given security change occurs. We use this approach to
control differences incurred by the specificities of each project. Still,
maintainability may be affected by different coding styles and experience level
from developers of the same project. We do not evaluate differences at that
level. Moreover, we do not evaluate the maintainbility difference for all
regular commits in a project. We assume that the size of the dataset is enough
to mitigate random variations in the maintainability differences of the baseline.

@@

The dataset used in this work comprises code changes of free and open
source applications. Our methodology requires access to data that is not
publicly available for commercial apps. The extent to which this findings
generalize to commercial applications with non-open source licenses remains
a question to be answered in future work.

Moreover, all the security commits in this work are described in English.
%
\section{Related Work}\label{sec:rw}

>>>>

\begin{abstract}
  Security is a crucial non-functionality requirement for software applications.
  However, building secure software is far from trivial as developers lack both
  the knowledge and tools to effectively address this concern. In this paper, we
  study the impact of changes to improve security on the maintainability of several
  open source applications. Using a dataset containing 607 security-oriented
  commits, we measure maintainability --- as computed by the Software Improvement
  Group's web-based source code analysis service \emph{Better Code Hub} (BCH) ---
  before and after the security refactoring. Results show that making software
  more secure comes at a cost on maintainability. This is particularly evident
  in refactorings to deal with \emph{Broken Authentication and Session Management}, \emph{Denial-of-Service} attacks and \emph{Memory Leaks}.
\end{abstract}

@@

The impact of each type of security refactoring on the maintainability of open-source software is presented in Figure \ref{fig:pat}. The results of the Wilcoxon signed-rank test by pattern returned statistical evidence which supports that the maintainability of open-source software decreases for 4 different patterns: Broken Autentication (p = 0.018), Denial-of-Service (p = 0.029), Memory Leak (p = 0.009) and Miscellaneous (p = $3.07$x$10^{-4}$). The results also returned statistical significance that evidence that maintainability is more likely to not suffer differences for open-source software after removing Components with Known Vulnerabilities (p = $7.96$x$10^{-4}$) and fixing Cross-Site Scripting Vulnerabilities (p = 0.001). Injection and Cross-Site Request Forgery produced more cases where the refactorings had a negative impact in the maintainability. However, the results obtained for these patterns are not statistically significant.

Projects with Broken Authentication flaws are the most affected ones by security refactorings. 55.3\% (21) of the refactorings had a negative impact in the maintainability while 35\% (11) had a positive impact and only 13.8\% (6) did not suffer any change. The results also show that fixing memory leaks affects the maintainability of 51.3\% (41) of the cases and leaves only 13.8\% (11) of the cases unaffected. 35\% (28) are affected positively. Refactoring vulnerabilities that allow denial-of-service attacks (e.g., CVE-2014-3506) has a negative impact of 47.5\% (19) in the maintainability of open-source software. 30.0\% (12) of the cases exhibit maintainability improvement while 22.5\% (9) show that overall maintainability does not suffer any change. Miscellaneous refactorings pattern integrates several types of security flaws whose impact is of 43.1\% (84 cases) negative in the maintainability of open-source sofware. 31.8\% (62) of the cases do not exhibit any change in maintainability and 25.1\% (49) of cases have a positive impact on the maintainability of open-source software.

It is also important to notice that overall the maintainability of open-source software is not affected by Cross-Site Scripting vulnerabilities and by Using Components with Know Vulnerabilities. Cross-Site Scripting does not suffer any decrease in the maintainability of 55.3\% (63) of cases and 21.1\% (24) have a positive effect on it. Only 23.7\% (27) of the cases hinder the software maintainability. From 21 cases evaluated for Using Componentes with Known Vulnerabilities, only 4.8\% (1) has a positive effect in the maintainability of software. 14.3\% (3) harms maintainability software and 81\% has no effect in the maintainability.

@@

In this section, we discuss the results and answer the proposed research questions.

@@

\textbf{\textit{Security refactorings harm the maintainability of open-source software.}}

@@

\textbf{\textit{Broken Authentication \& Session Management, Denial-of-Service attacks and Memory Leaks and Miscellaneous patterns are more likely to affect open-source software maintainability.}}

\section{Threats to Validity}\label{sec:threats}
%
The following section presents the potential threats to validity of this study.
%
\subsection{Construct}
%
The maintainability formula was inferred based on the reports of the Better Code Hub. 
Since the dataset has a broad group of open-source projects with different backgrounds, this metric may not be suitable for all of them. However, BCH does have a representative benchmark of
closed and open-source software projects that is under the calculation of the thresholds used in the computation of each maintainability guideline~\cite{Visser:2016:OREILLY, Baggen2012} (Figure \ref{fig:guidelines}).

@@

The security refactorings dataset provided by the previous
work~\cite{Reis:2017:IJSSE} was collected based on the messages of GitHub commits produced by
project developers to classify the changes perfomed by them to fix security flaws. This approach discards security refactorings that were not explicit in commits messages.

Baseline commits are retrieved randomly by selecting one commit from the same project of a given security refactoring. This approach softens the differences that may result from the
characteristics of each project. However, maintainability still may be affected by the developers experience, coding style and software contribution polices. But maintainability difference is not evaluated at this level.

@@

In contrast to the BCH benchmark, this study uses a dataset that only covers open-source software to perform the maintanibility evaluation. Thus, answering these questions to closed commercial applications may be a future work opportunity.

All the security commits are described in English.
%
\section{Related Work}\label{sec:rw}
